API design for machine learning software: experiences from the scikit-learn project

### 2013-08-09 paper.tex
Summarize discussion of #34

The library has been designed to tie in with the set of numeric
and scientific packages centered around the NumPy and SciPy libraries. NumPy
\citep{vanderwalt2011} augments Python with a contiguous numeric array datatype
and fast array computing primitives,
while SciPy \citep{varoquaux2013scipy} extends it
further with common numerical operations, either by implementing
these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
implementations. Building upon this stack, a series of libraries called
\textit{scikits} were created, to complement SciPy with
domain-specific toolkits. Currently, the two most popular and feature-complete
ones are by far \sklearn and
\textit{scikit-image},\footnote{\url{http://scikit-image.org}}
which does image processing.

Started in 2007, \sklearn is developed by an international team of over a dozen
core developers, mostly researchers from various fields (e.g.,
computer science, neuroscience, astrophysics). The project also benefits
from many occasional contributors proposing small bugfixes or
improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
a platform which greatly facilitates this kind of
collaboration. Because of the large number of developers, emphasis is
put on keeping the project maintainable. In particular, code must follow
specific quality guidelines, such as style consistency and unit-test coverage.
Documentation and examples are required for all features,
and major changes must pass code review by at least two
developers not involved in the implementation of the proposed change.

@@

\citet{pedregosa2011} briefly presented \sklearn and
benchmarked it against several competitors.
In this paper, we instead present an
in-depth analysis of design choices made when building the library,
detailing how we organized and operationalized
common machine learning concepts.
We first present in section~\ref{sec:core-api} the central application
programming interface (API) and then describe, in section~\ref{sec:advanced-api},
advanced API mechanisms built on the core interface.
Section~\ref{sec:implementation} briefly describes the implementation.
Section~\ref{sec:comparison} then
compares \sklearn to other major projects in terms of API\@.
Section~\ref{sec:future_work} outlines some of the objectives for
a \sklearn 1.0 release.
We conclude by summarizing the major points of this paper in
section~\ref{sec:conclusions}.

@@

Estimator initialization and actual learning are strictly separated,
in a way that is similar to partial function application:
an estimator is initialized from a set of named constant hyper-parameter values
(e.g., the $C$ constant in SVMs)
and can be considered as a function
that maps these values to actual learning algorithms.
The constructor of an estimator does not see any actual data, nor does it perform any actual learning.
All it does is attach the given parameters to the object.
For the sake of convenient model inspection, hyper-parameters are set as public attributes,
which is especially important in model selection settings.
For ease of use, default hyper-parameter values are also provided
for all built-in estimators.
These default values are set to be relevant in many common
situations in order to make estimators as effective as possible
\textit{out-of-box} for non-experts.

Actual learning is performed by the \texttt{fit} method. This method is called
with training data (e.g., supplied as two arrays \texttt{X\_train} and
\texttt{y\_train} in supervised learning estimators). Its task is to run a
learning algorithm and to determine model-specific parameters from the training
data and set these as attributes on the estimator object. As a convention, the
parameters learned by an estimator are exposed as public attributes with names
suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
learned coefficients of a linear model),
again to facilitate model inspection.
In the partial application view,
\texttt{fit} is a function from data to a model of that data.
As such, it always returns the estimator object it was called on,
which now serves as a model of its input and can be used to perform predictions or transformations of input data.
The choice to let a single object serve dual purpose as estimator and model
simplifies the \sklearn class hierarchy
compared to an approach where estimators and models are separated.

>>>>

The library has been designed to tie in with the set of numeric
and scientific packages centered around the NumPy and SciPy libraries. NumPy
\citep{vanderwalt2011} augments Python with a contiguous numeric array datatype
and fast array computing primitives,
while SciPy \citep{varoquaux2013scipy} extends it
further with common numerical operations, either by implementing
these in Python/NumPy or by wrapping existing C/C{}\verb!++!/Fortran
implementations. The \textit{scikits} were created to complement SciPy with
domain-specific toolkits. Currently, the two most popular and feature-complete
ones are by far \sklearn and
\textit{scikit-image},\footnote{\url{http://scikit-image.org}}
which does image processing.

Started in 2007, \sklearn is developed by an international team of a dozen of
core developers, mostly researchers from various fields (e.g.,
computer science, neuroscience, astrophysics). The project also benefits
from many occasional contributors proposing small bugfixes or
improvements. Development proceeds on GitHub\footnote{\url{https://github.com/scikit-learn}},
a platform which greatly facilitates this kind of
collaboration. Because of the large number of developers, emphasis is
put on keeping the project maintainable. In particular, code must follow
specific quality guidelines, such as style consistency and unit-test coverage.
Documentation and examples are required for all features,
and major changes must pass code review by at least two
developers not involved in the implementation of the proposed change.

@@

\citet{pedregosa2011} briefly presented \sklearn and
benchmarked it against several competitors.
In this paper, we instead present an
in-depth analysis of design choices made when building the library,
detailing how we organized and operationalized
common machine learning concepts.
We first present in section~\ref{sec:core-api} the central application
programming interface (API) and then describe, in section~\ref{sec:advanced-api},
advanced API mechanisms built on the core interface.
Section~\ref{sec:implementation} briefly describes the implementation.
Section~\ref{sec:comparison} then
compares \sklearn to other major projects in terms of API\@.
Section~\ref{sec:future_work} outlines some of the work to be done before
a \sklearn 1.0 release.
We summarize and conclude in section~\ref{sec:conclusions}.

@@

Estimator initialization and actual learning are strictly separated,
in a way that is similar to partial function application:
an estimator is initialized from a set of named constant hyper-parameter values
(e.g., the $C$ constant in SVMs)
and can be considered as a function
that maps these values to actual learning algorithms.
The constructor of an estimator does not see any actual data, nor does it perform any actual learning.
All it does is attach the given parameters to the object.
For the sake of convenient model inspection, hyper-parameters are set as public attributes,
which is especially important in model selection settings.
For ease of use, default hyper-parameter values are also provided
for all built-in estimators.
These default values are set to be relevant in most common
situations in order to make estimators as effective as possible
\textit{out-of-box} for non-experts.

Actual learning is performed by the \texttt{fit} method. This method is called
with training data (e.g., supplied as two arrays \texttt{X\_train} and
\texttt{y\_train} in supervised learning estimators). Its task is to run a
learning algorithm and to determine model-specific parameters from the training
data and set these as attributes on the estimator object. As a convention, the
parameters learned by an estimator are exposed as public attributes with names
suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
learned coefficients of a linear model),
again to facilitate model inspection.
In the partial application view,
\texttt{fit} is a function from data to a model of that data.
It always returns the estimator object it was called on,
which now serves as a model of its input and can be used to perform predictions or transformations of input data.

From the start, the choice to let a single object serve dual purpose as
estimator and model has mostly been driven by usability and technical
considerations. From the user point of view, having two coupled instances (i.e.,
an estimator object, used as a factory, and a model object, produced by the
estimator) indeed decreases the ease of use and is also more likely to
unecessarily confuse newcomers. From the developer point of view, decoupling
estimators from models also creates parallel class hierarchies and increases the
overall maintenance complexity of the project. For these practical reasons, we
believe that decoupling estimators from models is not worth the effort. The good
reason for decoupling however, would be that it makes it possible to ship a
model in a new environment without having to deal with potentially complex
software dependencies. Such a feature could however still be implemented in
\sklearn by making estimators able to export a fitted model to an agnostic model
description such as PMML.
